# Introduction to Statistical Learning - Chapter 2 #

## Question 8 ##
Read data from `College.csv`

```{r}
setwd("~/Dropbox/Code/R/IntroductionToStatisticalLearning")
df = read.csv(file = 'College.csv')
head(df)
```

Use column `X` as row names and remove it.

```{r}
rownames(df) = df[,1]
df$X = NULL
head(df)
```

Look at summary.

```{r}
summary(df)
```

Plot `Outstate` versus `Private`.

```{r}
library(ggplot2)
ggplot(data = df, mapping = aes(x=Outstate, y=Private)) + geom_point()
```

Create `Elite` variable.

```{r}
df$Elite = df$Top10perc>50
ggplot(data = df, mapping = aes(x=Outstate, y=Elite)) + geom_point()
```

## Question 9 ##

Read `Auto` data and remove rows with missing values.

```{r}
auto = read.csv(file = 'Auto.csv', header = T, na.strings = '?')
# look at rows containing NA
auto[apply(is.na(auto), MARGIN = 1, FUN = any),]
# remove rows containing NA
auto = na.omit(auto)
```

`mpg, displacement, horsepower, weight` and `acceleration` are quantitative variables. Rest are qualitative. Convert qualitative variables to factors.

```{r}
auto$cylinders = as.factor(auto$cylinders)
auto$year = as.factor(auto$year)
auto$origin = as.factor(auto$origin)
# name is already a factor
# look at ranges of quantitative variables
range(auto$mpg)
range(auto$displacement)
range(auto$horsepower)
range(auto$weight)
range(auto$acceleration)
```

Look at summary. 

```{r}
summary(auto)
```

Remove observations 10 to 85, and look at the summary.

```{r}
summary(auto[-c(10:85),])
```

Look at pairwise scatterplots.

```{r}
pairs(auto)
```

## Question 10 ##

Load `Boston` dataset from `MASS` package.

```{r}
library(MASS)
head(Boston)
```

How many rows and columns?

```{r}
dim(Boston)
```

Look at some pairwise scatterplots of crime rate.

```{r}
par(mfrow = c(2,2))
plot(Boston$medv, Boston$crim)
plot(Boston$rm, Boston$crim)
plot(Boston$ptratio, Boston$crim)
plot(Boston$age, Boston$crim)
```

Some suburbs have unusually high crime rates.

```{r}
summary(Boston$crim)
hist(Boston$crim, breaks = 50)
```

Look at suburbs with low median value of homes.

```{r}
head(Boston[order(Boston$medv),], 20)
```

Look at suburbs with high average number of rooms per dwelling.

```{r}
head(Boston[order(Boston$rm, decreasing = T),], 20)
```










